Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2021-06-04 doc/manuscript/ESSDmanuscript.Rmd
misc updates

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'ESSD')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "2" 
  - given_name: Ben 
    surname: Bond-Lamberty (confirm)
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Sandro 
    surname: Federici 
    affiliation: "5" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "6" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "7" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Valentyna
    surname: Slivinska 
    affiliation: "5"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,6 *"
    email: teixeirak@si.edu
    corresponding: true

@@

Here, we:
(1) clarify definitions of relevant carbon stocks and increments
(2) describe mapping of ForC to IPCC's EFDB,
(3) describe updates to ForC  (ForC v4.0),
(4) summarize the data in ForC that's relevant to EFDB, identifying gaps, and
(5) provide recommendations for improving data collection, analysis, database, and accounting.
We note that

@@

To support export of data to EFDB, and to improve the overall quality of the ForC database, we implemented some modest restructuring, resolved duplicate records, and conducted quality control.

This section describes changes relative to ForC v2.0 [@anderson-teixeira_forc_2018].

@@

**Table: changes to ForC fields**.  
Use  this table : https://github.com/forc-db/ForC/blob/master/database_management_records/record_of_changes.csv.

@@

## Quality control

@@

summarize the data in ForC that's relevant to EFDB, identifying gaps

@@

**A universal challenge in estimating biomass (living or dead) from forest census data is applying appropriate allometries to convert DBH measurements to biomass.**
*(Camille/Helene can write this paragraph easily.)*

@@

**We recommend that, unless they have some specific reason to do otherwise, researchers calculate and report the values according to IPCC standards:**

@@

@@

>>>>

---
title: Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'ESSD')`"
author:
  - given_name: Madison 
    surname: Williams
    affiliation: "1"
  - given_name: Valentine
    surname: Herrmann
    affiliation: "1"
  - given_name: Rebecca
    surname: Banbury Morgan
    affiliation: "1,2" 
  - given_name: Ben 
    surname: Bond-Lamberty (confirm)
    affiliation: "3" 
  - given_name: Susan 
    surname: Cook-Patton
    affiliation: "4" 
  - given_name: Sandro 
    surname: Federici 
    affiliation: "5" 
  - given_name: Helene
    surname: Muller-Landau
    affiliation: "6" 
  - given_name: Camille
    surname: Piponiot
    affiliation: "7" 
  - given_name: Teagan
    surname: Rogers
    affiliation: "1"
  - given_name: Valentyna
    surname: Slivinska 
    affiliation: "5"
  - given_name: Kristina J.
    surname: Anderson-Teixeira
    affiliation: "1,6 *"
    email: teixeirak@si.edu
    corresponding: true

@@

Here, we:
(1) clarify definitions of relevant carbon stocks and increments
(2) describe mapping of ForC to IPCC's EFDB,
(3) describe updates to ForC  (ForC v4.0),
(4) summarize the data in ForC that's relevant to EFDB, identifying gaps, and
(5) provide recommendations for improving data collection, analysis, database, and accounting.

@@

To support export of data to EFDB, and to improve the overall quality of the ForC database, we implemented some modest restructuring, resolved duplicate records, and conducted quality control.
This section describes changes relative to ForC v2.0 [@anderson-teixeira_forc_2018].

@@

```{r, out.width = "12cm", echo = FALSE,fig.pos="H", fig.cap = "Table of changes to ForC fields (placeholder) "}
#this is for a two-column figure. for one-column figure, use {r, out.width = "8.3cm", echo = FALSE, fig.cap = "one column figure"}
knitr::include_graphics("figures_tables/ForC_updates_placeholder_fig.png")
```
*(The above is a placeholder for the table located at  https://github.com/forc-db/ForC/blob/master/database_management_records/record_of_changes.csv, which we'll need to format.)*

## Quality control measures
Prior to releasing ForC v4.0, we executed several quality control measures.
First, to improve information on geographic coordinates, we flagged and reviewed records with suspected low precision *(Issue #29)[https://github.com/forc-db/ForC/issues/229]*.
Second, to identify erroneous climate data... *(Issue #212)[https://github.com/forc-db/ForC/issues/212]*.

@@

@@

**(summarize the data in ForC that's relevant to EFDB, identifying gaps)**

@@

A universal challenge in estimating biomass (living or dead) from forest census data is applying appropriate allometries to convert DBH measurements to biomass.
*(Camille/Helene can write this paragraph easily.)*

@@

We recommend that, unless they have some specific reason to do otherwise, researchers calculate and report the values according to IPCC standards:

@@

**Contributing data to ForC and/or EFDB directly will ensure its broader impact.**
The latter is more efficient for getting data to EFDB, but does not get the data into ForC, where it can be more broadly useful--for example, being used for basic science [e.g., @banburymorgan_global_2021; @anderson-teixeira_carbon_2021] or model benchmarking [@fer_ecosystem_2021].

@@

Remote sensing biomass estimates include standing dead wood [@duncanson_aboveground_2021].

### 2021-06-04 doc/manuscript/ESSDmanuscript.tex
misc updates

\Author[1]{Madison}{Williams}
\Author[1]{Valentine}{Herrmann}
\Author[2]{Rebecca}{Banbury Morgan}
\Author[3]{Ben}{Bond-Lamberty (confirm)}
\Author[4]{Susan}{Cook-Patton}
\Author[5]{Sandro}{Federici}
\Author[6]{Helene}{Muller-Landau}
\Author[7]{Camille}{Piponiot}
\Author[1]{Teagan}{Rogers}
\Author[5]{Valentyna}{Slivinska}
\Author[1,6 *]{Kristina J.}{Anderson-Teixeira}

@@

To support export of data to EFDB, and to improve the overall quality of
the ForC database, we implemented some modest restructuring, resolved
duplicate records, and conducted quality control.

This section describes changes relative to ForC v2.0
\citep{anderson-teixeira_forc_2018}.

@@

\textbf{Table: changes to ForC fields}.\\
Use this table :
https://github.com/forc-db/ForC/blob/master/database\_management\_records/record\_of\_changes.csv.

@@

\subsection{Quality control}

@@

summarize the data in ForC that's relevant to EFDB, identifying gaps

@@

\textbf{A universal challenge in estimating biomass (living or dead)
from forest census data is applying appropriate allometries to convert
DBH measurements to biomass.} \emph{(Camille/Helene can write this
paragraph easily.)}

@@

\textbf{We recommend that, unless they have some specific reason to do
otherwise, researchers calculate and report the values according to IPCC
standards:}

@@

@@

>>>>

\Author[1]{Madison}{Williams}
\Author[1]{Valentine}{Herrmann}
\Author[1,2]{Rebecca}{Banbury Morgan}
\Author[3]{Ben}{Bond-Lamberty (confirm)}
\Author[4]{Susan}{Cook-Patton}
\Author[5]{Sandro}{Federici}
\Author[6]{Helene}{Muller-Landau}
\Author[7]{Camille}{Piponiot}
\Author[1]{Teagan}{Rogers}
\Author[5]{Valentyna}{Slivinska}
\Author[1,6 *]{Kristina J.}{Anderson-Teixeira}

@@

To support export of data to EFDB, and to improve the overall quality of
the ForC database, we implemented some modest restructuring, resolved
duplicate records, and conducted quality control. This section describes
changes relative to ForC v2.0 \citep{anderson-teixeira_forc_2018}.

@@

\begin{figure}[H]
\includegraphics[width=12cm]{figures_tables/ForC_updates_placeholder_fig} \caption{Table of changes to ForC fields (placeholder) }\label{fig:unnamed-chunk-2}
\end{figure}

\emph{(The above is a placeholder for the table located at
https://github.com/forc-db/ForC/blob/master/database\_management\_records/record\_of\_changes.csv,
which we'll need to format.)}

\subsection{Quality control measures}

Prior to releasing ForC v4.0, we executed several quality control
measures. First, to improve information on geographic coordinates, we
flagged and reviewed records with suspected low precision \emph{(Issue
\#29){[}https://github.com/forc-db/ForC/issues/229{]}}. Second, to
identify erroneous climate data\ldots{} \emph{(Issue
\#212){[}https://github.com/forc-db/ForC/issues/212{]}}.

@@

@@

\textbf{(summarize the data in ForC that's relevant to EFDB, identifying
gaps)}

@@

A universal challenge in estimating biomass (living or dead) from forest
census data is applying appropriate allometries to convert DBH
measurements to biomass. \emph{(Camille/Helene can write this paragraph
easily.)}

@@

We recommend that, unless they have some specific reason to do
otherwise, researchers calculate and report the values according to IPCC
standards:

@@

\textbf{Contributing data to ForC and/or EFDB directly will ensure its
broader impact.} The latter is more efficient for getting data to EFDB,
but does not get the data into ForC, where it can be more broadly
useful--for example, being used for basic science
\citep[e.g.,][]{banburymorgan_global_2021, anderson-teixeira_carbon_2021}
or model benchmarking \citep{fer_ecosystem_2021}.

@@

Remote sensing biomass estimates include standing dead wood
\citep{duncanson_aboveground_2021}.
